---
title: "Constitute 9-proteins' HCC proteomic subtypes identifing model"
institute:
  - Mengchao Hepatobiliry Hospital
author:
  - En Hu
keywords:
  - subtype
  - HCC 
  - model
documentclass: ctexart
geometry: margin=0.6cm
output:
  rticles::ctex:
    fig_caption: yes
    number_sections: yes
    toc: yes
header-includes:
 - \usepackage{fvextra}
 - \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
 - \CTEXsetup[format={\huge\bfseries\raggedright}]{section}
---

# Environment

```{r}
rm(list = ls())
options(stringsAsFactors = F)
Sys.setlocale(locale = "English")
pacman::p_load(magrittr,tidyr,stringr,dplyr,DMwR2,DreamAI,factoextra,FactoMineR,ggalluvial,ggforce,caret,Boruta,parallel,do,pROC,ggpubr)
windowsFonts(TNR=windowsFont("Times new Roman"))
source("Function/Function.R")
```

-   Here, three cohorts' proteomic data was used to constitute subtype identifing model.(Jiang et als.' data, Gao et als.' data and This study's data).

-   Only tumor sample's expression data was used in model building.

# 1. Input data

## 1.1 Jiang et als.' data

```{r}
load("./1_raw_data/2. Jiang data/jiang_raw_expression.Rdata")
load("./1_raw_data/2. Jiang data/jiang_sur_data.Rdata")
```

## 1.2 Gao et als.' data

```{r}
load("./1_raw_data/3. Gao data/gao_raw_expression.Rdata")
load("./1_raw_data/3. Gao data/gao_sur_data.Rdata")
```

## 1.3 This study's data

```{r}
load("./1_raw_data/1.Proteomic data/152 samples and 4032 proteins .Rdata")
load("./2_output_data/sur_data_152.data")
load("./1_raw_data/1.Proteomic data/dec.Rdata")
```

# 2. Data Cleaning

## 2.1 Missing value imputation for Jiang's data

```{r}
clean_jiang_exp <- jiang_raw_expression[apply(jiang_raw_expression[99:199],1,function(l) sum(!is.na(l))>101*0.4),#proteins identified more than 40% tumor samples
                       99:199]   %>% as.matrix() %>%
  impute.RegImpute(.,fillmethod = "row_mean",maxiter_RegImpute = 10,conv_nrmse = 1e-06) %>%
                            2^. %>%
  as.data.frame(.) %>% 
  apply(.,2,function(x) x/median(x)) %>%
  log2() %>%
  as.data.frame()
 
```

## 2.2 Intersection of three cohots's data

```{r}
clean_gao_exp <- gao_raw_expression[1:159]
clean_this_exp <- tumor_exp

intersection_proteins <- dec$PG.Genes %>% is_in(dec$PG.ProteinGroups %>% is_in(intersect(rownames(clean_this_exp),rownames(clean_jiang_exp)) ) %>%
  which() %>% dec$PG.Genes[.] %>% intersect(rownames(clean_gao_exp))) %>% which() %>%
  dec$PG.ProteinGroups[.]

intersection_genes <- dec$PG.ProteinGroups %>% is_in(intersection_proteins) %>%
  which() %>% dec$PG.Genes[.]

jiang_exp_prepared_1 <- clean_jiang_exp[intersection_proteins,]
gao_exp_prepared_1 <- clean_gao_exp[intersection_genes,] %>% set_rownames(intersection_proteins)
this_exp_prepared_1 <- clean_this_exp[intersection_proteins,]
```

## 2.3 Prognostic data of three cohors

```{r}
all.sur.data <- rbind(sur_data_152[c("OS.time","OS","RFS.time","RFS")],
                      jiang_sur_data[c("OS.time","OS","RFS.time","RFS")],
                      gao_sur_data[c("OS.time","OS","RFS.time","RFS")]) %>%
                mutate(subtype =
                     c(as.character(sur_data_152$sub761),
                      jiang_sur_data$`subtype_by_Liu_sig(721)`,
                      gao_sur_data$`subtype_by_gao_sig(1274)`))
```

## 2.4 Scale expression data

There is obvious batch effect in three cohort due to difference of data output platform. Then, all data were scaled for eliminating it.

### 2.4.1 Data distribution before removal of batch effects

```{r}
all_batch_data <- cbind(this_exp_prepared_1,jiang_exp_prepared_1,gao_exp_prepared_1)
p2.4.1 <- Plot_pca(PCA(t(all_batch_data)),group = c(rep("This cohort",152),rep("Jiang's cohort",101),
                                     rep("Gao's cohort",159)))
p2.4.1
```

### 2.4.2 Data distribution afte removal of batch effects

```{r}
all_no_batch_data <- cbind(this_exp_prepared_1 %>% t() %>% scale() %>% t() %>% as.data.frame(),
                           jiang_exp_prepared_1%>% t() %>% scale() %>% t() %>% as.data.frame(),
                           gao_exp_prepared_1%>% t() %>% scale() %>% t() %>% as.data.frame())
p2.4.2 <- Plot_pca(PCA(t(all_no_batch_data)),group = c(rep("This cohort",152),rep("Jiang's cohort",101),
                                     rep("Gao's cohort",159)))
p2.4.2
```

# 3. Partitioning the training and validation sets

## 3.1 training

```{r}
train_data <- rbind(t(this_exp_prepared_1 %>% t() %>% scale() %>% t() %>% as.data.frame()),
                    t(jiang_exp_prepared_1%>% t() %>% scale() %>% t() %>% as.data.frame())) %>% 
                    as.data.frame()
train_group <- all.sur.data[rownames(train_data),"subtype"] 
colnames(train_data) <- dec$PG.Genes[match(colnames(train_data),dec$PG.ProteinGroups)] %>%
  gsub("-","_",.)
train_data[train_data >2] <- 2
train_data[train_data < -2] <- -2
```

## 3.2 validation

```{r}
var_data<- t(gao_exp_prepared_1%>% t() %>% scale() %>% t() %>% as.data.frame()) %>% as.data.frame()
var_group <- all.sur.data[rownames(var_data),"subtype"]
colnames(var_data) <- dec$PG.Genes[match(colnames(var_data),dec$PG.ProteinGroups)] %>%
  gsub("-","_",.)
var_data[var_data < -2] <- -2
var_data[var_data < -2] <- -2
```

# 4. Feature section

## 4.1 comparison between any two of three subtypes

```{r}

train.sub.diff <- lapply(rownames(t(train_data)),function(i){
  x = t(train_data)[i,]
  SIvsSII.count = 0
  SIvsSIII.count = 0
  SIIvsSIII.count = 0
  SIvsSII.FC = c()
  SIIvsSIII.FC =c()
  SIvsSIII.FC=c()
  #Repeat Random Sampling 50 times(80% samples per time)
  for(i in 1:50){
    
    index = createDataPartition(train_group,p = 0.8 )

    exp = t(x)[index$Resample1]
    
    sample = rownames(train_data)[index$Resample1]

    s1 = which(sample %in% rownames(all.sur.data)[all.sur.data$subtype=="SI"])
    s2 = which(sample %in% rownames(all.sur.data)[all.sur.data$subtype=="SII"])
    s3 = which(sample %in% rownames(all.sur.data)[all.sur.data$subtype=="SIII"])

    SIvsSII.p = wilcox.test(exp[s1],exp[s2])[[3]];
    SIIvsSIII.p = wilcox.test(exp[s3],exp[s2])[[3]];
    SIvsSIII.p = wilcox.test(exp[s1],exp[s3])[[3]];

    SIvsSII.fc = 2^(mean(exp[s1]) - mean(exp[s2]))
    SIIvsSIII.fc = 2^(mean(exp[s2]) - mean(exp[s3]))
    SIvsSIII.fc = 2^(mean(exp[s1]) - mean(exp[s3]))

    if(SIvsSII.p<0.01&(SIvsSII.fc>log2(1.5)|SIvsSII.fc<log2(1/1.5))){
      SIvsSII.count = SIvsSII.count + 1
      SIvsSII.FC = c(SIvsSII.FC,SIvsSII.fc)
    }
    if(SIIvsSIII.p<0.01&(SIIvsSIII.fc>log2(1.5)|SIIvsSIII.fc<log2(1/1.5))){
      SIIvsSIII.count = SIIvsSIII.count + 1
      SIIvsSIII.FC = c(SIIvsSIII.FC,SIIvsSIII.fc)
    }
    if(SIvsSIII.p<0.01&(SIvsSIII.fc>log2(1.5)|SIvsSIII.fc<log2(1/1.5))){
      SIvsSIII.count = SIvsSIII.count + 1
      SIvsSIII.FC = c(SIvsSIII.FC,SIvsSIII.fc)
    }
  }

  if(length(SIvsSII.FC)>0){
    SIvsSII.FC.p = wilcox.test(SIvsSII.FC)[[3]]
    SIvsSII.FC.m = mean(SIvsSII.FC)
  }else{
    SIvsSII.FC.p = 1
    SIvsSII.FC.m = 0
  }
  if(length(SIIvsSIII.FC)>0){
    SIIvsSIII.FC.p = wilcox.test(SIIvsSIII.FC)[[3]]
    SIIvsSIII.FC.m = mean(SIIvsSIII.FC)
  }else{
    SIIvsSIII.FC.p = 1
    SIIvsSIII.FC.m = 0
  }
  if(length(SIvsSIII.FC)>0){
    SIvsSIII.FC.p = wilcox.test(SIvsSIII.FC)[[3]]
    SIvsSIII.FC.m = mean(SIvsSIII.FC)
  }else{
    SIvsSIII.FC.p = 1
    SIvsSIII.FC.m = 0
  }

  return(data.frame(SIvsSII.count = SIvsSII.count,
                    SIIvsSIII.count = SIIvsSIII.count,
                    SIvsSIII.count = SIvsSIII.count,
                    SIvsSII.FC.p =  SIvsSII.FC.p,
                    SIvsSII.FC.m = SIvsSII.FC.m ,
                    SIIvsSIII.FC.p =  SIIvsSIII.FC.p ,
                    SIIvsSIII.FC.m = SIIvsSIII.FC.m ,
                    SIvsSIII.FC.p =SIvsSIII.FC.p,
                    SIvsSIII.FC.m = SIvsSIII.FC.m
  )

  )
}) %>%  do.call(rbind,.)
```

## 4.2 comparison between any one subtype and the other two subtypes

```{r}
train.sub.diff2 <- lapply(colnames(train_data),function(i){
  x = t(t(train_data)[i,])
  
  SI.count = 0
  SIII.count = 0
  SII.count = 0
  SI.FC = c()
  SIII.FC =c()
  SII.FC=c()
  
  for(i in 1:50){
    
    index = createDataPartition(train_group,p = 0.8 )
    
    exp = t(x)[index$Resample1]
    sample = rownames(train_data)[index$Resample1]
   
    s1 = which(sample %in% rownames(all.sur.data)[all.sur.data$subtype=="SI"])
    s2 = which(sample %in% rownames(all.sur.data)[all.sur.data$subtype=="SII"])
    s3 = which(sample %in% rownames(all.sur.data)[all.sur.data$subtype=="SIII"])

    SI.p = wilcox.test(exp[s1],exp[s2])[[3]];
    SII.p = wilcox.test(exp[s3],exp[s2])[[3]];
    SIII.p = wilcox.test(exp[s1],exp[s3])[[3]];

    SI.fc = 2^(mean(exp[s1]) - mean(exp[s2]))
    SII.fc = 2^(mean(exp[s2]) - mean(exp[s3]))
    SIII.fc = 2^(mean(exp[s1]) - mean(exp[s3]))

    if(SI.p<0.01&(SI.fc>log2(1.5)|SI.fc<log2(1/1.5))){
      SI.count = SI.count + 1
      SI.FC = c(SI.FC,SI.fc)
    }
    if(SII.p<0.01&(SII.fc>log2(1.5)|SII.fc<log2(1/1.5))){
      SII.count = SII.count + 1
      SII.FC = c(SII.FC,SII.fc)
    }
    if(SIII.p<0.01&(SIII.fc>log2(1.5)|SIII.fc<log2(1/1.5))){
      SIII.count = SIII.count + 1
      SIII.FC = c(SIII.FC,SIII.fc)
    }
  }
  
  if(length(SI.FC)>0){
    SI.FC.p = wilcox.test(SI.FC)[[3]]
    SI.FC.m = mean(SI.FC)
  }else{
    SI.FC.p = 1
    SI.FC.m = 0
  }
  if(length(SII.FC)>0){
    SII.FC.p = wilcox.test(SII.FC)[[3]]
    SII.FC.m = mean(SII.FC)
  }else{
    SII.FC.p = 1
    SII.FC.m = 0
  }
  if(length(SIII.FC)>0){
    SIII.FC.p = wilcox.test(SIII.FC)[[3]]
    SIII.FC.m = mean(SIII.FC)
  }else{
    SIII.FC.p = 1
    SIII.FC.m = 0
  }

  return(data.frame(SI.count = SI.count,
                    SII.count = SII.count,
                    SIII.count = SIII.count,
                    SI.FC.p =  SI.FC.p,
                    SI.FC.m = SI.FC.m ,
                    SII.FC.p =  SII.FC.p ,
                    SII.FC.m = SII.FC.m ,
                    SIII.FC.p = SIII.FC.p,
                    SIII.FC.m = SIII.FC.m
  )
  )
}) %>%  do.call(rbind,.)
rownames(train.sub.diff) <- colnames(train_data)
rownames(train.sub.diff2) <- colnames(train_data)
```

## 4.3 intersection

```{r}
#comparison between any two of three subtypes
sig1 <-  train.sub.diff %>% filter(.,SIvsSII.count>40& (abs(log2(SIvsSII.FC.m))>0.58)|
                                     SIIvsSIII.count>40&(abs(log2(SIIvsSIII.FC.m))>0.58)|
                                     SIIvsSIII.count>40&(abs(log2(SIvsSIII.FC.m))>0.58)
                                     )
#comparison between any one subtype and the other two subtypes
sig2 <-  train.sub.diff2 %>% filter(.,SI.count>40& (abs(log2(SI.FC.m))>0.58)|
                                     SII.count>40&(abs(log2(SII.FC.m))>0.58)|
                                     SIII.count>40&(abs(log2(SIII.FC.m))>0.58)
                                     )
sig3 <- intersect(rownames(sig1),rownames(sig2))      


```

## 4.4 remove hihg correlation proteins

```{r}
train_sig3_exp <- train_data[sig3]
descrCorr = cor(train_sig3_exp)
highCorr = findCorrelation(descrCorr, 0.9)
train_sig4_exp = train_sig3_exp[, -highCorr]
colnames(train_sig4_exp) <- gsub("-","_",colnames(train_sig4_exp))

```

## 4.5 Boruta

### 4.5.1 Boruta for (SI vs others)

```{r}
boruta.data  <-  train_sig4_exp %>%
  mutate(subtype=as.factor(gsub("SIII","SII",train_group )))
boru.rf <- Boruta(subtype~.,data=boruta.data, pValue = 0.01, mcAdj = TRUE, doTrace = 2)
boru.rf.result <- boru.rf$finalDecision %>% as.data.frame()
boru.rf.confirm_SI <- rownames(boru.rf.result)[boru.rf.result$.=="Confirmed"]
```

### 4.5.2 Boruta for (SII vs others)

```{r}

boruta.data  <-  train_sig4_exp %>%
  mutate(subtype=as.factor(gsub("^SII$","SI",train_group )))
boru.rf <- Boruta(subtype~.,data=boruta.data, pValue = 0.01, mcAdj = TRUE, doTrace = 2)
boru.rf.result <- boru.rf$finalDecision %>% as.data.frame()
boru.rf.confirm_SIII <- rownames(boru.rf.result)[boru.rf.result$.=="Confirmed"]
```

# 5 knn model for SIvsOthers

```{r}
control <- trainControl(method="cv",5,summaryFunction=twoClassSummary,classProbs = TRUE)
```

## 5.1 Get Importance Ranking

```{r}
knn_SI_1 <- train_data[boru.rf.confirm_SI] %>%
  mutate(subtype=as.factor(gsub("SIII","SII",train_group_2)))
knn_SI_result1 <- train(subtype~.,knn_SI_1,
              method = 'knn',
              trControl = control,
              tuneLength = 30,
              metric="ROC",
)

knn_SI_result1_im <- varImp(knn_SI_result1)[[1]]
knn_SI_result1_im <- rownames(knn_SI_result1_im)[order(knn_SI_result1_im$SI,decreasing = T)]
```

## 5.2 Forward selection method based on importance ranking

```{r}
cl <- makeCluster(30)
clusterExport(cl,varlist = c("train_data","knn_SI_result1_im","train_group","control"))
knn_SI_result2 <- parLapply(cl,1:30,function(i){
  pacman::p_load("caret","do","magrittr","dplyr")
  message(paste0("第",i,"个蛋白"))
  #i=1
  t.data <- train_data[knn_SI_result1_im[c(1:i)]] %>%
    mutate(subtype=as.factor(gsub("SIII","SII",train_group)))
  t.summary <- lapply(1:30,function(x){
    
    t.model <- train(subtype~.,t.data,
                     method = 'knn',
                     #preProcess = c('center','scale'),
                     trControl = control,
                     tuneLength = 30,
                     metric="ROC"
                     
    )
    
    
    
    return(data.frame(
      AUC = t.model[["resample"]][1] %>% t() %>% mean(),
      Sens = t.model[["resample"]][2] %>% t() %>% mean(),
      Spec = t.model[["resample"]][3] %>% t() %>% mean()
    ))
  }) %>% do.call(rbind,.) %>% t() %>% as.data.frame()
  
  t.summary$M <- apply(t.summary,1,mean)
  t.summary$sd <- apply(t.summary,1,sd)
  return(t.summary)
})
stopCluster(cl)

# 绘制蛋白累积折线图
knn_SI_accumulate <- do.call(rbind,knn_SI_result2)[31:32]
knn_SI_accumulate$tag <- rep(c("AUC","Sens","Spec"))
knn_SI_accumulate$tag2 <- rep(c(1:30),each=3)
knn_SI_accumulate$tag3 <- ifelse((knn_SI_accumulate$tag %in% c("Sens","Spec")) &
                                 (knn_SI_accumulate$tag2 %in% c("4","5")),"Rejected","Accepeted"
)
p5.2 <- ggplot(data = knn_SI_accumulate,aes(x=tag2,y=M))+
  
  geom_segment(x=20,xend=20,y=0.65,yend=1,color="red",size=0.8,linetype=2)+
  geom_errorbar(aes(ymin=M-sd,ymax=M+sd,width=.1),color="grey40")+
  geom_line(aes(group=tag,color=tag),size=1)+
  geom_point(aes(shape=tag3,color=tag3))+
  
  scale_x_continuous(breaks = c(1:30))+
  scale_y_continuous(breaks = seq(0.5,1,0.025),limits = c(0.65,1))+
  theme_classic()+
  scale_color_manual(breaks = c("Rejected","Accepeted","AUC","Sens","Spec"),
                     values = c("red","black","#845EC2","#936C00","#008CCC"))+
  labs(x="Number of proteins used in knn model",y="",color="")+
  theme(text = element_text(family = "TNR",face="bold",color = "black",size = 10),
        axis.title  = element_text(size = 15),
        legend.text = element_text(size = 13))

cairo_pdf(filename = "./3_output_figure/subtype_panel_SIvsOthers_forwardsection_1.pdf",width = 6,height = 5)
p5.2
dev.off()

```

Through the "p5.2" picture, we found there was no more significant change in model accuracy from the fourth proteins, but the sensitivity fluctuated more. In particular, the third and fourth proteins caused a substantial decrease in sensitivity, so we removed these two proteins in the next model training.

## 5.3 Forward selection method based on importance ranking_2

```{r}
knn_SI_result1_im_2  <- knn_SI_result1_im[-4:-5]
cl <- makeCluster(30)
clusterExport(cl,varlist = c("train_data","knn_SI_result1_im_2","train_group","control"))
knn_SI_result3 <- parLapply(cl,c(1:30),function(i){
  pacman::p_load("caret","do","magrittr","dplyr")
  #i=1
  t.data <- train_data[knn_SI_result1_im_2[c(1:i)]] %>%
    mutate(subtype=as.factor(gsub("SIII","SII",train_group)))
  t.summary <- lapply(1:30,function(x){
    
    t.model <- train(subtype~.,t.data,
                     method = 'knn',
                     #preProcess = c('center','scale'),
                     trControl = control,
                     tuneLength = 30,
                     metric="ROC"
                     
    )
    
    
    
    return(data.frame(
      AUC = t.model[["resample"]][1] %>% t() %>% mean(),
      Sens = t.model[["resample"]][2] %>% t() %>% mean(),
      Spec = t.model[["resample"]][3] %>% t() %>% mean()
    ))
  }) %>% do.call(rbind,.) %>% t() %>% as.data.frame()
  
  t.summary$M <- apply(t.summary,1,mean)
  t.summary$sd <- apply(t.summary,1,sd)
  return(t.summary)
})
stopCluster(cl)
knn_SI_accumulate2 <- do.call(rbind,knn_SI_result3)[31:32]
knn_SI_accumulate2$tag <- rep(c("AUC","Sens","Spec"))
knn_SI_accumulate2$tag2 <- rep(c(1:30),each=3)
knn_SI_accumulate2$tag3 <- ifelse((knn_SI_accumulate2$tag %in% c("Sens","Spec")) &
                                 (knn_SI_accumulate2$tag2 %in% "4"),"Rejected","Accepeted"
)
p5.3 <- ggplot(data = knn_SI_accumulate2,aes(x=tag2,y=M))+
  
  geom_segment(x=18,xend=18,y=0.65,yend=1,color="red",size=0.8,linetype=2)+
  geom_errorbar(aes(ymin=M-sd,ymax=M+sd,width=.1),color="grey30")+
  geom_line(aes(group=tag,color=tag),size=1)+
  geom_point(aes(color=tag3,shape=tag3))+
  
  scale_x_continuous(breaks = c(1:30))+
  scale_y_continuous(breaks = seq(0.5,1,0.025),limits = c(0.65,1))+
  theme_classic()+
  scale_color_manual(breaks = c("Rejected","Accepeted","AUC","Sens","Spec"),
                     values = c("red","black","#845EC2","#936C00","#008CCC"))+
  labs(x="Number of proteins used in knn model",y="",color="")+
  theme(text = element_text(family = "TNR",face="bold",color = "black",size = 10),
        axis.title  = element_text(size = 15),
        legend.text = element_text(size = 13))

cairo_pdf(filename = "./3_output_figure/subtype_panel_SIvsOthers_forwardsection_2.pdf",width = 6,height = 5)
p5.3
dev.off()
```

In a new round of training, we found that the 4th protein (formerly the 6th protein) led to a decrease in sensitivity.So the protein 6 was also rejected.

## 5.4 Forward selection method based on importance ranking_3

```{r}
knn_SI_result1_im_3  <- knn_SI_result1_im[-4:-6]
cl <- makeCluster(30)
clusterExport(cl,varlist = c("train_data","knn_SI_result1_im_3","train_group","control"))
knn_SI_result4 <- parLapply(cl,c(1:30),function(i){
  pacman::p_load("caret","do","magrittr","dplyr")
  #i=1
  t.data <- train_data[knn_SI_result1_im_3[c(1:i)]] %>%
    mutate(subtype=as.factor(gsub("SIII","SII",train_group)))
  t.summary <- lapply(1:30,function(x){
    
    t.model <- train(subtype~.,t.data,
                     method = 'knn',
                     #preProcess = c('center','scale'),
                     trControl = control,
                     tuneLength = 30,
                     metric="ROC"
                     
    )
    
    
    
    return(data.frame(
      AUC = t.model[["resample"]][1] %>% t() %>% mean(),
      Sens = t.model[["resample"]][2] %>% t() %>% mean(),
      Spec = t.model[["resample"]][3] %>% t() %>% mean()
    ))
  }) %>% do.call(rbind,.) %>% t() %>% as.data.frame()
  
  t.summary$M <- apply(t.summary,1,mean)
  t.summary$sd <- apply(t.summary,1,sd)
  return(t.summary)
})
stopCluster(cl)
knn_SI_accumulate3 <- do.call(rbind,knn_SI_result4)[31:32]
knn_SI_accumulate3$tag <- rep(c("AUC","Sens","Spec"))
knn_SI_accumulate3$tag2 <- rep(c(1:30),each=3)

p5.4 <- ggplot(data = knn_SI_accumulate3,aes(x=tag2,y=M))+
  
  geom_segment(x=4,xend=4,y=0.65,yend=1,color="red",size=0.8,linetype=2)+
  geom_errorbar(aes(ymin=M-sd,ymax=M+sd,width=.1),color="grey30")+
  geom_line(aes(group=tag,color=tag),size=1)+
  geom_point()+
  
  scale_x_continuous(breaks = c(1:30))+
  scale_y_continuous(breaks = seq(0.5,1,0.025),limits = c(0.65,1))+
  theme_classic()+
  scale_color_manual(values = c("#845EC2","#936C00","#008CCC"))+
  labs(x="Number of proteins used in knn model",y="",color="")+
  theme(text = element_text(family = "Times New Roman",face="bold",color = "black",size = 10),
        axis.title  = element_text(size = 15),
        legend.text = element_text(size = 13))

cairo_pdf(filename = "./3_output_figure/subtype_panel_SIvsOthers_forwardsection_3.pdf",width = 6,height = 5)
p5.4
dev.off()
```

Ultimately, we found that the performance of the models built using the first four and the seventh protein were more balanced 
##  5.5 the end

```{r}
knn_SI_end <- train_data[knn_SI_result1_im[c(1:3,7)]] %>%
  mutate(subtype=as.factor(gsub("SIII","SII",train_group)))
set.seed(1)
knn_SI_result_end <- train(subtype~.,knn_SI_end ,
              method = 'knn',
              #preProcess = c('center','scale'),
              trControl = control,
              tuneLength = 30,
              metric="ROC"
              
)
predict(knn_SI_result_end,var_data) %>% confusionMatrix(.,as.factor(gsub("SIII","SII",var_group)))
```

 Confusion Matrix and Statistics

               Accuracy : 0.8868         
                 95% CI : (0.827, 0.9315)
    No Information Rate : 0.6478         
    P-Value [Acc > NIR] : 5.661e-12      
                                         
                  Kappa : 0.7478         
                                         
Mcnemar's Test P-Value : 0.4795

            Sensitivity : 0.8036         
            Specificity : 0.9320         
         Pos Pred Value : 0.8654         
         Neg Pred Value : 0.8972         
             Prevalence : 0.3522         
         Detection Rate : 0.2830         

Detection Prevalence : 0.3270\
Balanced Accuracy : 0.8678

       'Positive' Class : SI   

# 6 knn model for SIIIvsOthers

## 6.1 Get Importance Ranking

```{r}
knn_SIII_1 <- train_data[boru.rf.confirm_SIII] %>%
  mutate(subtype=as.factor(gsub("^SII$","SI",train_group)))
knn_SIII_result1 <- train(subtype~.,knn_SIII_1,
              method = 'knn',
              trControl = control,
              tuneLength = 30,
              metric="ROC",
)

knn_SIII_result1_im <- varImp(knn_SIII_result1)[[1]]
knn_SIII_result1_im <- rownames(knn_SIII_result1_im)[order(knn_SIII_result1_im$SI,decreasing = T)]
```

## 6.2 Forward selection method based on importance ranking

```{r}
cl <- makeCluster(30)
clusterExport(cl,varlist = c("train_data","knn_SIII_result1_im","train_group","control"))
knn_SIII_result2 <- parLapply(cl,1:30,function(i){
  pacman::p_load("caret","do","magrittr","dplyr")
  #i=1
  t.data <- train_data[knn_SIII_result1_im[c(1:i)]] %>%
    mutate(subtype=as.factor(gsub("^SII$","SI",train_group)))
  t.summary <- lapply(1:30,function(x){
    
    t.model <- train(subtype~.,t.data,
                     method = 'knn',
                     #preProcess = c('center','scale'),
                     trControl = control,
                     tuneLength = 30,
                     metric="ROC"
                     
    )
    
    
    
    return(data.frame(
      AUC = t.model[["resample"]][1] %>% t() %>% mean(),
      Sens = t.model[["resample"]][2] %>% t() %>% mean(),
      Spec = t.model[["resample"]][3] %>% t() %>% mean()
    ))
  }) %>% do.call(rbind,.) %>% t() %>% as.data.frame()
  
  t.summary$M <- apply(t.summary,1,mean)
  t.summary$sd <- apply(t.summary,1,sd)
  return(t.summary)
})
stopCluster(cl)

# 绘制蛋白累积折线图
knn_SIII_accumulate <- do.call(rbind,knn_SIII_result2)[31:32]
knn_SIII_accumulate$tag <- rep(c("AUC","Sens","Spec"))
knn_SIII_accumulate$tag2 <- rep(c(1:30),each=3)
knn_SIII_accumulate$tag3 <- ifelse((knn_SIII_accumulate$tag %in% c("Sens","Spec")) &
                                 (knn_SIII_accumulate$tag2 %in% c("3","4","5","6","7","10","12")),"Rejected","Accepeted"
)

p6.2 <- ggplot(data = knn_SIII_accumulate,aes(x=tag2,y=M))+
  
  geom_segment(x=13,xend=13,y=0.75,yend=1,color="red",size=0.8,linetype=2)+
  geom_errorbar(aes(ymin=M-sd,ymax=M+sd,width=.1),color="grey30")+
  geom_line(aes(group=tag,color=tag),size=1)+
  geom_point(aes(shape=tag3,color=tag3))+
  
  scale_x_continuous(breaks = c(1:30))+
  scale_y_continuous(breaks = seq(0.5,1,0.025),limits = c(0.7,1))+
  theme_classic()+
  scale_color_manual(breaks = c("Rejected","Accepeted","AUC","Sens","Spec"),
                     values = c("red","black","#845EC2","#936C00","#008CCC"))+
  labs(x="Number of proteins used in knn model",y="",color="")+
  theme(text = element_text(family = "TNR",face="bold",color = "black",size = 10),
        axis.title  = element_text(size = 15),
        legend.text = element_text(size = 13))

cairo_pdf(filename = "./3_output_figure/subtype_panel_SIIIvsOthers_forwardsection_1.pdf",width = 6,height = 5)
p6.2
dev.off()
```
Through the picture "p6.2", we found there was no more significant change in model accuracy from the fourth proteins, but the sensitivity fluctuated more. In particular, the (3,4,5,6,7,10,12) proteins caused a substantial decrease in sensitivity, so we removed these seven proteins in the next model training.

## 6.3 Forward selection method based on importance ranking_2

```{r}
cl <- makeCluster(30)
knn_SIII_result1_im_2 <- knn_SIII_result1_im[c(-3:-7,-10,-12)]
clusterExport(cl,varlist = c("train_data","knn_SIII_result1_im_2","train_group","control"))
knn_SIII_result3 <- parLapply(cl,1:30,function(i){
  pacman::p_load("caret","do","magrittr","dplyr")
  #i=1
  t.data <- train_data[knn_SIII_result1_im_2[c(1:i)]] %>%
    mutate(subtype=as.factor(gsub("^SII$","SI",train_group)))
  t.summary <- lapply(1:30,function(x){
    
    t.model <- train(subtype~.,t.data,
                     method = 'knn',
                     #preProcess = c('center','scale'),
                     trControl = control,
                     tuneLength = 30,
                     metric="ROC"
                     
    )
    
    
    
    return(data.frame(
      AUC = t.model[["resample"]][1] %>% t() %>% mean(),
      Sens = t.model[["resample"]][2] %>% t() %>% mean(),
      Spec = t.model[["resample"]][3] %>% t() %>% mean()
    ))
  }) %>% do.call(rbind,.) %>% t() %>% as.data.frame()
  
  t.summary$M <- apply(t.summary,1,mean)
  t.summary$sd <- apply(t.summary,1,sd)
  return(t.summary)
})
stopCluster(cl)

# 绘制蛋白累积折线图
knn_SIII_accumulate2 <- do.call(rbind,knn_SIII_result3)[31:32]
knn_SIII_accumulate2$tag <- rep(c("AUC","Sens","Spec"))
knn_SIII_accumulate2$tag2 <- rep(c(1:30),each=3)


p6.3 <- ggplot(data = knn_SIII_accumulate2, aes(x=tag2,y=M))+
  
  geom_segment(x=6,xend=6,y=0.75,yend=1,color="red",size=0.8,linetype=2)+
  geom_errorbar(aes(ymin=M-sd,ymax=M+sd,width=.1),color="grey30")+
  geom_line(aes(group=tag,color=tag),size=1)+
  geom_point()+
  
  scale_x_continuous(breaks = c(1:30))+
  scale_y_continuous(breaks = seq(0.5,1,0.025),limits = c(0.7,1))+
  theme_classic()+
  scale_color_manual(values = c("#845EC2","#936C00","#008CCC"))+
  labs(x="Number of proteins used in knn model",y="",color="")+
  theme(text = element_text(family = "Times New Roman",face="bold",color = "black",size = 10),
        axis.title  = element_text(size = 15),
        legend.text = element_text(size = 13))

cairo_pdf(filename = "./3_output_figure/subtype_panel_SIIIvsOthers_forwardsection_2.pdf",width = 6,height = 5)
p6.3
dev.off()
```
1,2,8,9,11,13 proteins were selected
## 6.4 the end


```{r}
knn_SIII_end <- train_data[knn_SIII_result1_im[c(1:2,8,9,11,13)]] %>%
  mutate(subtype=as.factor(gsub("^SII$","SI",train_group)))
set.seed(29)
knn_SIII_result_end <- train(subtype~.,knn_SIII_end ,
              method = 'knn',
              #preProcess = c('center','scale'),
              trControl = control,
              tuneLength = 30,
              metric="ROC"
              
)

predict(knn_SIII_result_end,var_data) %>% confusionMatrix(.,as.factor(gsub("^SII$","SI",var_group)))
```
Confusion Matrix and Statistics
                                          
               Accuracy : 0.956           
                 95% CI : (0.9114, 0.9821)
    No Information Rate : 0.6981          
    P-Value [Acc > NIR] : <2e-16          
                                          
                  Kappa : 0.8974          
                                          
 Mcnemar's Test P-Value : 0.4497          
                                          
            Sensitivity : 0.9550          
            Specificity : 0.9583          
         Pos Pred Value : 0.9815          
         Neg Pred Value : 0.9020          
             Prevalence : 0.6981          
         Detection Rate : 0.6667          
   Detection Prevalence : 0.6792          
      Balanced Accuracy : 0.9566          
                                          
       'Positive' Class : SI 
# 7. Integration of two models
```{r}
SI_SII.predict = predict(knn_SI_result_end,var_data) %>% gsub("^SII$","SII+SIII",.) %>%
  as.data.frame() %>% set_rownames(rownames(var_data))
SI_SIII.predict = predict(knn_SIII_result_end,var_data) %>% gsub("^SI$","SI+SII",.) %>%
  as.data.frame() %>% set_rownames(rownames(var_data))

end <- cbind(SI_SII.predict,SI_SIII.predict)
colnames(end) <- c("m1","m2")

end$end <- ifelse(end$m1=="SI"&end$m2=="SI+SII","SI",
                  ifelse(end$m1=="SII+SIII"&end$m2=="SIII","SIII","SII"))
end$sub <- var_group

caret::confusionMatrix(factor(end$sub),
                       factor(end$end))
```
## 7.1 Figure2G Muti-ROC curve 
```{r}
end$SI <- end$sub %>% gsub("SIII","SII",.)
end$SI.p <- end$end %>% gsub("SIII","SII",.)
end$SIII <- end$sub %>% gsub("^SII$","SI",.)
end$SIII.p <- end$end %>% gsub("^SII$","SI",.)
end$SII <- end$sub %>% gsub("SIII","SI",.)
end$SII.p <- end$end %>% gsub("SIII","SI",.)
roc1 <- roc(end$SI%>% as.factor()  ,end$SI.p %>% as.factor() %>% as.numeric() )
roc2 <- roc(end$SIII%>% as.factor()  ,end$SIII.p %>% as.factor() %>% as.numeric() )
roc3 <- roc(end$SII%>% as.factor()  ,end$SII.p %>% as.factor() %>% as.numeric() )
multiclass.roc(end$sub %>% as.factor(),end$end %>% as.factor() %>% as.numeric()) -> mulirc

 mul.roc.p <- ggroc(list(roc1,roc3,roc2))+
  geom_line(aes(color=name),size=1.2)+
  scale_color_manual(values = c("green","blue","red"),label=c("SI vs SII&SIII","SII vs SI&SIII","SIII vs SI&SII"))+
  theme_bw()+
  theme(#axis.line = element_line(size=1),
        plot.background = element_blank(),
        panel.grid = element_blank(),
        panel.border = element_rect(size=1.3,color="black"),
        axis.text = element_text(family = "TNR",color = "black",size=12),
        axis.title = element_text(family = "TNR",size = 15),
        legend.title = element_blank(),
        legend.position = "top",
        legend.text = element_text(family = "TNR",size=10)
        )+
  geom_segment(x=1,xend=-1,y=2,yend=0,linetype=2,color="grey",size=0.8)+
  geom_text(aes(x=0.3,y=0.8,label="Multi-class area under the curve: 0.925"))
 
 auc.table <- data.frame(
  SI=c(0.865,0.897,0.867),
  SII=c(0.758,0.891,0.832),
  SIII = c(0.918,0.972,0.950),
  row.names = c("Sensitivity","Specificity","AUC"))
roc.table <- ggtexttable(auc.table,theme = ttheme("light",rownames.style = rownames_style(face="bold")))%>%
                table_cell_bg(.,row=4,column = 4,fill="red")


cairo_pdf(filename = "./3_output_figure/F2G_1.pdf",width = 6,height = 5)
 mul.roc.p
dev.off()

cairo_pdf(filename = "./3_output_figure/F2G_2.pdf",width = 6,height = 5)
roc.table
dev.off()
```
## 7.2 Figure2H sankey plot
```{r}
sankey.data <- data.frame(Cohort = rep(c("Response","Predictor"),each=159),
                          response = c(end$sub,end$end) %>% factor(.,levels = c("SIII","SII","SI")),
                          subject = rep(c(1:159),2),
                          freq = 1)

sankey.plot <- ggplot(sankey.data,
       aes(x = Cohort, stratum = response%>% factor(.,levels = c("SIII","SII","SI")),
           alluvium = subject,
           y = freq,
           fill = response, label = response)) +
  scale_x_discrete(expand = c(.1, .1)) +
  geom_flow(color="black",size=0.8) +
  geom_stratum(alpha = .8,size=1) +
  geom_text(stat = "stratum", size = 4)+
  coord_flip()+
  theme(panel.background = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank())+
  scale_fill_manual(values = c("#FB050B","#0081FF","#009944"))
cairo_pdf(filename = "./3_output_figure/F2H.pdf",width = 6,height = 3)
sankey.plot
dev.off()

```
# save

```{r}
save(list = ls(),file = "./2_output_data/2.Constitute 9-proteins' HCC proteomic subtypes identifing model.Rdata")

```
