---
title: "Code and code of Figure 1 "
author: 
  - HuEn
  - "`r Sys.Date()`"
  
documentclass: ctexart
geometry: margin=0.6cm
output:
  rticles::ctex:
    fig_caption: yes
    number_sections: yes
    toc: yes
header-includes:
 - \usepackage{fvextra}
 - \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
 - \CTEXsetup[format={\huge\bfseries\raggedright}]{section}
---

# 1. Load data
```{r}
options(stringsAsFactors = F)
pacman::p_load(magrittr, dplyr, ggpubr,aplot,clusterProfiler,GSVA,factoextra,ggalluvial)
source("Function/Function.R")
load("./1_raw_data/1.Proteomic data/sur152.Rdata")
load("1_raw_data/1.Proteomic data/sur_data_30.Rdata")
load("./1_raw_data/1.Proteomic data/sur_data_70.Rdata")
load("1_raw_data/1.Proteomic data/152 samples and 4032 proteins .Rdata")
load("1_raw_data/1.Proteomic data/152 samples and all proteins.Rdata")
load("2_output_data/NA_stastistic.Rdata")
load("1_raw_data/1.Proteomic data/train set clinicopathoglogic.Rdata")
kegg_class <- read.csv("./1_raw_data/1.Proteomic data/2_kegg_class.csv")
```
# 2. data set
```{r}
train_set <- dataT[rownames(sur_data_70)]

valid_set <- dataT[rownames(sur_data_30)]

```
#3. subtype of train set
## 3.1 feature section
```{r}
## median absolute deviation
mad_train_set_1500 <- rownames(train_set[order(apply(train_set,1,mad), decreasing = T)[1:1500],])
## proteins identified in more than 80% samples
proteins_80 <- rownames(pro_6512)[apply(pro_6512[colnames(train_set)],1,function(x) sum(!is.na(x)))>=108*0.8]
feature <- intersect(mad_train_set_1500,proteins_80)
```
## 3.2 consensuscluster
```{r}
train_result <- ConsensusClusterPlus(as.matrix(train_set[feature,]), maxK = 8,
                                                           reps = 1000, pItem = 0.8,
                                                           pFeature = 0.8,  
                                                           clusterAlg = 'kmdist', 
                                                           distance = 'euclidean',
                                                           title ="./3_output_figure/liu_sig1128_70percentsample",
                                                           plot = "pdf"
                                                            )

sub_group <- data.frame(train_result[[4]]$consensusClass)

sur_data_70$train_set_1128 <- sub_group[,1]
fit <- survfit(Surv(OS.time, OS) ~ train_set_1128, sur_data_70)
ggsurvplot(fit, sur_data_70,pval = TRUE)
#绘制热图
sur_data_70$train_set_1128 <- sur_data_70$train_set_1128 %>%
  gsub("1","SI",.) %>% 
  gsub("3","SII",.) %>%
  gsub("2","SIII",.) %>%
  gsub("4", "SIII",.)
```
## 3.3 subtype-specfic proteins
```{r}
SI = which(colnames(train_set) %in% rownames(sur_data_70)[sur_data_70$train_set_1128=="SI"])
SII = which(colnames(train_set) %in% rownames(sur_data_70)[sur_data_70$train_set_1128=="SII"])
SIII = which(colnames(train_set) %in% rownames(sur_data_70)[sur_data_70$train_set_1128=="SIII"])
subtype_dep <- apply(train_set,1,function(p){
  data.frame(SIIvsSI.p = wilcox.test(p[SI],p[SII])[[3]],
             SIIvsSI.FC = mean(2^p[SII])/mean(2^p[SI]),
             SIIIvsSII.p = wilcox.test(p[SII],p[SIII])[[3]],
             SIIIvsSII.FC = mean(2^p[SIII])/mean(2^p[SII]),
             SIIIvsSI.p = wilcox.test(p[SIII],p[SI])[[3]],
             SIIIvsSI.FC = mean(2^p[SIII])/mean(2^p[SI])
             )
}) %>% do.call(rbind,.)

sub_spec <- filter(subtype_dep,(SIIvsSI.p<0.01&(SIIvsSI.FC >2 |SIIvsSI.FC <0.5))|
                                  (SIIIvsSII.p<0.01&(SIIIvsSII.FC >2 |SIIIvsSII.FC <0.5))|
                                 (SIIIvsSI.p<0.01&(SIIIvsSI.FC >2 |SIIIvsSI.FC <0.5)))
sub_specfics <- intersect(rownames(sub_spec),proteins_80)
```
## 3.4 subtype with sub_specfic proteins(761)
```{r}
train_result2 <- ConsensusClusterPlus(as.matrix(train_set[sub_specfics,]), maxK = 8,
                                                           reps = 1000, pItem = 0.8,
                                                           pFeature = 0.8,  
                                                           clusterAlg = 'kmdist', 
                                                           distance = 'euclidean',
                                                           title ="./3_output_figure/liu_sig761_70percentsample",
                                                           plot = "pdf"
                                                            )

sub_group2 <- data.frame(train_result2[[4]]$consensusClass)

sur_data_70$train_set_761 <- sub_group[,1]
fit <- survfit(Surv(OS.time, OS) ~ train_set_761, sur_data_70)
ggsurvplot(fit, sur_data_70,pval = TRUE)
#绘制热图
sur_data_70$train_set_761 <- sur_data_70$train_set_761 %>%
  gsub("1","SI",.) %>% 
  gsub("3","SII",.) %>%
  gsub("2","SIII",.) %>%
  gsub("4", "SIII",.)
```
# 4.subtype of valid_set
```{r}
valid_result <- ConsensusClusterPlus(as.matrix(valid_set[feature,]), maxK = 8,
                                                           reps = 1000, pItem = 0.8,
                                                           pFeature = 0.8,  
                                                           clusterAlg = 'kmdist', 
                                                           distance = 'euclidean',
                                                           title ="./3_output_figure/liu_sig1128_30percentsample",
                                                           plot = "pdf"
                                                            )

sub_group3 <- data.frame(valid_result[[4]]$consensusClass)

sur_data_30$valid_set_1128 <- sub_group3[,1]
fit <- survfit(Surv(OS.time, OS) ~ valid_set_1128, sur_data_30)
ggsurvplot(fit, sur_data_30,pval = TRUE)
#绘制热图
sur_data_30$valid_set_1128 <- sur_data_30$valid_set_1128 %>%
  gsub("1","SI",.) %>% 
  gsub("2","SII",.) %>%
  gsub("3","SIII",.) %>%
  gsub("4", "SIII",.)
sur_data_30$valid_set_1128 == sur_data_30$Subtype152sample_3
```
## 4.1 subtype of valid_set with 761 proteins
```{r}
valid_result2 <- ConsensusClusterPlus(as.matrix(valid_set[sub_specfics,]), maxK = 8,
                                                           reps = 1000, pItem = 0.8,
                                                           pFeature = 0.8,  
                                                           clusterAlg = 'kmdist', 
                                                           distance = 'euclidean',
                                                           title ="./3_output_figure/liu_sig761_30percentsample",
                                                           plot = "pdf"
                                                            )

sub_group4 <- data.frame(valid_result2[[4]]$consensusClass)

sur_data_30$valid_set_761 <- sub_group4[,1]
fit <- survfit(Surv(OS.time, OS) ~ valid_set_761, sur_data_30)
ggsurvplot(fit, sur_data_30,pval = TRUE)
#绘制热图
sur_data_30$valid_set_761 <- sur_data_30$valid_set_761 %>%
  gsub("1","SI",.) %>% 
  gsub("2","SII",.) %>%
  gsub("3","SIII",.) %>%
  gsub("4", "SIII",.)
sur_data_30$valid_set_761 == sur_data_30$Subtype_30sample_761sig_3sub
table(sur_data_30$valid_set_761,sur_data_30$valid_set_1128)
```
## 4.2 152 samples’ subtype
```{r}
sur_152.data$sub761 <- c(sur_data_70[rownames(sur_152.data),"train_set_761"] %>% na.omit(),
                         sur_data_30[rownames(sur_152.data),"valid_set_761"] %>% na.omit())

```
# 5. heatmp of train set
## 5.1 pheatmap
```{r}
p51.anno <- data.frame(sample=rownames(sur_data_70),
                       sub=sur_data_70$train_set_761)

## sample_order
SI = which(colnames(train_set) %in% rownames(sur_data_70)[sur_data_70$train_set_1128=="SI"])
SII = which(colnames(train_set) %in% rownames(sur_data_70)[sur_data_70$train_set_1128=="SII"])
SIII = which(colnames(train_set) %in% rownames(sur_data_70)[sur_data_70$train_set_1128=="SIII"])
train_set_order <- c(SI,SII,SIII) %>% colnames(train_set)[.]


## annotation
ggplot(p51.anno,aes(x=sample,y="",fill=sub))+
  geom_tile()+
  theme_void()+
  scale_fill_manual(values = c("forestgreen","midnightblue","red4"))->p51.1

## expression
p51.exp <- train_set[sub_specfics,] %>%  t() %>% scale() %>% t() %>% as.data.frame()
p51.exp[p51.exp < -2] <- -2; p51.exp[p51.exp > 2 ]<- 2

## gene order
gene_order <- dist(p51.exp) %>% hclust() %>%  use_series(order) %>% rownames(p51.exp)[.]

p51.2.plot.data <- p51.exp %>% mutate(gene=rownames(p51.exp)) %>%
  reshape2::melt(id.vars="gene") %>% 
  mutate(sample = factor(variable,levels = train_set_order)) %>%
  mutate(gene = factor(gene,levels = gene_order)) %>%
  mutate(sub = sur_data_70[variable,"train_set_761"])

## plot
ggplot(p51.2.plot.data,aes(x=sample,y=gene,fill=value))+
  geom_tile()+
  theme_bw()+
  scale_fill_gradientn(colours = colorRampPalette(c("forestgreen", "grey0", "red2"))(200))+
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        panel.grid = element_blank(),
        axis.ticks = element_blank())+
  labs(fill="Expression")->p51.2

p51.2 %>% insert_top(p51.1,height = 0.05) -> p51.12
```
## 5.2 Clinicopathoglogic anno
```{r} 

anno_AFP <- data.frame(row.names = rownames(phe70_fix),
                       "AFP\n **" = phe70_fix$AFP
)
anno_MVI <- data.frame(row.names = rownames(phe70_fix),
                       "MVI\n *" = phe70_fix$MVI
)
anno_Differentiation <- data.frame(row.names = rownames(phe70_fix),
                                   "Differentiation\n **" = phe70_fix$Differentiation)
anno_TNM <- data.frame(row.names = row.names(phe70_fix),
                       TNM = phe70_fix$TNM.stage)
anno_BCLC <- data.frame(row.names = rownames(phe70_fix),
                        BCLC = phe70_fix$BCLC.stage)
anno_AGE <- data.frame(row.names = rownames(phe70_fix),
                       Age = phe70_fix$Age)
anno_Gender <- data.frame(row.names = rownames(phe70_fix),
                          Gender = phe70_fix$Gender)
anno_capp <- data.frame(row.names = rownames(phe70_fix),
                        "Tumor capsule" = phe70_fix$Tumor.capsule)
anno_number <- data.frame(row.names = rownames(phe70_fix),
                        "Tumor number" = phe70_fix$Tumor.number)
anno_size <- data.frame(row.names = rownames(phe70_fix),
                          "Tumor size" = phe70_fix$Tumor.size)
anno_PVTT <- data.frame(row.names = rownames(phe70_fix),
                        "PVTT"= phe70_fix$PVTT) 

#---------------------------------------------------------
p.AFP <- F.clinico.anno(sample.anno = anno_AFP,anno = "AFP",sample.order = train_set_order)+
  scale_fill_manual(values = c("#424642","#c06014"))+
  theme(axis.text.y = element_text(color = "red"))+
  scale_y_discrete(labels=c("AFP ***"),position = "right")
p.MVI <- F.clinico.anno(sample.anno = anno_MVI,anno = "MVI",sample.order = train_set_order)+
  scale_fill_manual(values = c("#FDE8CD","#00917c"))+
  theme(axis.text.y = element_text(color="red"))+
  scale_y_discrete(labels=c("MVI *"),position = "right")
p.TNM <- F.clinico.anno(sample.anno = anno_TNM,anno = "TNM",sample.order = train_set_order)+
  scale_fill_manual(breaks=c("I","II","III","IV"),values = c("#E7D9EA","#16C79A","#11689E","#19456B"))

p.BCLC <- F.clinico.anno(sample.anno = anno_BCLC,anno = "BCLC",sample.order = train_set_order)+
  scale_fill_manual(values = c("#FFFF66","#FFCC00","#FF9900","#FF0000"))
p.diff <- F.clinico.anno(sample.anno = anno_Differentiation,
                        anno="Differentiation",sample.order = train_set_order)+
  scale_fill_manual(breaks=c("high","medium","low","No"),values = c("#F2F4C3","#FC92E3","#005792","#480032"),
                    labels= c("High","Medium","Low","No"))+
  theme(axis.text.y =element_text(color="red"))+
  scale_y_discrete(labels=c("Differentiation ***"),position = "right")
p.TNM <- F.clinico.anno(sample.anno = anno_TNM,
                        anno="TNM stage",sample.order = train_set_order)+
  scale_fill_manual(values = c("#75D4F2","#52BDF2","#2D7DBC","#00416D"))
p.BCLC <- F.clinico.anno(sample.anno = anno_BCLC,
                        anno="BCLC",sample.order = train_set_order)+
  scale_fill_manual(values = c("#E1E7EB","#AEB8C3","#947B89","#4E0E2E"))

p.age<- F.clinico.anno(sample.anno = anno_AGE,
                        anno="Age",sample.order = train_set_order)+
  scale_fill_manual(breaks = c("<40","40~60",">60"),values = c("#D25565","#2E94B9","#F0B775"))
p.gender<- F.clinico.anno(sample.anno = anno_Gender,
                      anno="Gender",sample.order = train_set_order)+
  scale_fill_manual(values = c("#FF82A9","#7F95D1"))
p.capp<- F.clinico.anno(sample.anno = anno_capp,
                         anno="Tumor capsule",sample.order = train_set_order)+
  scale_fill_manual(values = c("#248888","#E7475E","#F0D879"),
                    breaks = c("complete","Incomplete","No"),
                    labels=c("Complete","Incomplete","No"))
p.number <- F.clinico.anno(sample.anno = anno_number,
                          anno="Tumor-number",sample.order = train_set_order)+
  scale_fill_manual(values = c("#f9eed3","#9fb083"),labels=c("Mutilple","Solitary"))
p.size <- F.clinico.anno(sample.anno = anno_size,
                          anno="Tumor-size",sample.order = train_set_order)+
  scale_fill_manual(values = c("#de4307","#f6d04d","#8bc24c"),
                    breaks =c("<5","5~10",">10"))
p.pvtt <- F.clinico.anno(sample.anno = anno_PVTT,
                        anno="PVTT",sample.order = train_set_order)+
  scale_fill_manual(values = c("#808180FF","#3B4992FF"))
#---------------------------------------------------------
p51.1 %>% 
  #insert_top(p.sub,height = .02) %>%
  insert_top(p.AFP,height = .02) %>%
  insert_top(p.MVI,height = .02) %>%
  insert_top(p.diff,height = .02) %>%
  insert_top(p.TNM,height = .02) %>%
  insert_top(p.BCLC,height = .02) %>%
  insert_top(p.age,height = .02) %>%
  insert_top(p.gender,height = .02) %>%
  insert_top(p.capp,height = .02) %>%
  insert_top(p.number,height = .02) %>%
  insert_top(p.size,height = .02) %>%
  insert_top(p.pvtt,height = .02)-> p51.3
```
## 5.3 save figure1A(figure1A)
```{r}
cairo_pdf(filename = "./3_output_figure/F1A_1.pdf",width = 7,height = 15)
print(p51.3)
dev.off()
cairo_pdf(filename = "./3_output_figure/F1A_2.pdf",width = 7,height = 7)
print(p51.12)
dev.off()
#---------------------------------------------------------
p51.3.pot.data <- reshape2::melt(anno_clinico,id.vars="sample")

ggplot(p51.3.pot.data,aes(x=sample,y="",fill=value))+
  geom_tile()+
  theme_void()
```
# 6. K-M curve(Figure1B,Figure1C)
## 6.1 train set
```{r}
## OS
legend_label <- data.frame(table(sur_data_70$train_set_761)) %$% 
  paste0(Var1,": ","n=",Freq)
fit <- survfit(Surv(OS.time,OS)~train_set_761,sur_data_70)
Cairo::CairoPDF(file = "./3_output_figure/F1B_1.pdf",width = 6,height = 5)
F.plot_surival(fit_object = fit,
             surivaldata = sur_data_70, 
             ytitle = OS, legend_text =  legend_label,palettes = c("forestgreen","midnightblue","red4")
)
dev.off()
## RFS
legend_label <- data.frame(table(sur_data_70$train_set_761)) %$% 
  paste0(Var1,": ","n=",Freq)
fit <- survfit(Surv(RFS.time,RFS)~train_set_761,sur_data_70)
Cairo::CairoPDF(file = "./3_output_figure/F1B_2.pdf",width = 6,height = 5)
F.plot_surival(fit_object = fit,
             surivaldata = sur_data_70, 
             ytitle = RFS, legend_text =  legend_label,palettes = c("forestgreen","midnightblue","red4")
)
dev.off()

```
## 6.2 valid_set
```{r}
legend_label <- data.frame(table(sur_data_30$valid_set_761)) %$% 
  paste0(Var1,": ","n=",Freq)
fit <- survfit(Surv(OS.time,OS)~valid_set_761,sur_data_30)
Cairo::CairoPDF(file = "./3_output_figure/F1C_1.pdf",width = 6,height = 5)
F.plot_surival(fit_object = fit,
             surivaldata = sur_data_30, 
             ytitle = OS, legend_text =  legend_label,palettes = c("forestgreen","midnightblue","red4")
)
dev.off()
## RFS
legend_label <- data.frame(table(sur_data_30$valid_set_761)) %$% 
  paste0(Var1,": ","n=",Freq)
fit <- survfit(Surv(RFS.time,RFS)~valid_set_761,sur_data_30)
Cairo::CairoPDF(file = "./3_output_figure/F1C_2.pdf",width = 6,height = 5)
F.plot_surival(fit_object = fit,
             surivaldata = sur_data_30, 
             ytitle = RFS, legend_text =  legend_label,palettes = c("forestgreen","midnightblue","red4")
)
dev.off()

## combine SI with SII
sur_data_30$valid_set_761_2 <- sur_data_30$valid_set_761 %>% gsub("^SI$","SI&SII",.) %>% gsub("^SII$","SI&SII",.)

legend_label <- data.frame(table(sur_data_30$valid_set_761_2)) %$% 
  paste0(Var1,": ","n=",Freq)
fit <- survfit(Surv(RFS.time,RFS)~valid_set_761_2,sur_data_30)
cairo_pdf(file = "./3_output_figure/F1C_2_2.pdf",width = 6,height = 5)
F.plot_surival(fit_object = fit,
             surivaldata = sur_data_30, 
             ytitle = RFS, legend_text =  legend_label,palettes = c("forestgreen","midnightblue","red4")
)
dev.off()


```
# 7. HR(figure1B、figure1C)
## 7.1 train_set
```{r}
#OS
########## calculate
F.pasteHR <- function(conf.int){
  paste0(round(conf.int[1],2),"(",round(conf.int[3],2),"-",round(conf.int[4],2),")")
}
F.pasteHR(SIIvsSI$conf.int)
SIIvsSI = coxph(Surv(OS.time,OS)~train_set_761,sur_data_70[sur_data_70$train_set_761!="SIII",]) %>% summary()
SIIIvsSI = coxph(Surv(OS.time,OS)~train_set_761,sur_data_70[sur_data_70$train_set_761!="SII",]) %>% summary()
SIIIvsSII = coxph(Surv(OS.time,OS)~train_set_761,sur_data_70[sur_data_70$train_set_761!="SI",]) %>% summary()

#########plot

data.frame(#subtypes = c("SI","SII","SIII"),
           SI = c("",SIIvsSI$logtest[3]%>% round(.,4),SIIIvsSI$logtest[3]%>% round(.,4)),
           SII = c(F.pasteHR(SIIvsSI$conf.int),"",SIIIvsSII$logtest[3]%>% round(.,4)),
           SIII = c(F.pasteHR(SIIIvsSI$conf.int),F.pasteHR(SIIIvsSII$conf.int),"")
           ) %>% 
ggtexttable(rows = c("SI","SII","SIII"),cols=c("SI","SII","SIII"),
            theme = ttheme(colnames.style=colnames_style(color = "black",face = "bold",size = 12,fill = "grey80",linewidth = 1,
  linecolor = "white"),rownames.style=rownames_style(  color = "black",
  face = "bold",
  size = 12,
  fill = "grey80",
  linewidth = 1,
  linecolor = "white"))) %>%
 table_cell_bg(row = c(3),column = c(2),fill= ggplot2::alpha("#79AF97FF",0.6)) %>%
      table_cell_bg(row = c(4),column = c(2),fill=ggplot2::alpha("#79AF97FF",0.6)) %>%
      table_cell_bg(row = c(4),column = c(3),fill=ggplot2::alpha("#79AF97FF",0.6))  %>%
      table_cell_bg(row = c(2),column = c(3),fill=ggplot2::alpha("#B24745FF",0.6)) %>%
      table_cell_bg(row = c(2),column = c(4),fill=ggplot2::alpha("#B24745FF",0.6)) %>%
      table_cell_bg(row = c(3),column = c(4),fill=ggplot2::alpha("#B24745FF",0.6))-> x
cairo_pdf(file = "./3_output_figure/F1B_1_HR.pdf",width = 6,height = 5)
x$theme$text$family="TNR"
print(x)
dev.off()

#RFS
########## calculate

SIIvsSI = coxph(Surv(RFS.time,RFS)~train_set_761,sur_data_70[sur_data_70$train_set_761!="SIII",]) %>% summary()
SIIIvsSI = coxph(Surv(RFS.time,RFS)~train_set_761,sur_data_70[sur_data_70$train_set_761!="SII",]) %>% summary()
SIIIvsSII = coxph(Surv(RFS.time,RFS)~train_set_761,sur_data_70[sur_data_70$train_set_761!="SI",]) %>% summary()

#########plot

data.frame(#subtypes = c("SI","SII","SIII"),
           SI = c("",SIIvsSI$logtest[3]%>% round(.,4),SIIIvsSI$logtest[3]%>% round(.,4)),
           SII = c(F.pasteHR(SIIvsSI$conf.int),"",SIIIvsSII$logtest[3]%>% round(.,4)),
           SIII = c(F.pasteHR(SIIIvsSI$conf.int),F.pasteHR(SIIIvsSII$conf.int),"")
           ) %>% 
ggtexttable(rows = c("SI","SII","SIII"),cols=c("SI","SII","SIII"),
            theme = ttheme(colnames.style=colnames_style(color = "black",face = "bold",size = 12,fill = "grey80",linewidth = 1,
  linecolor = "white"),rownames.style=rownames_style(  color = "black",
  face = "bold",
  size = 12,
  fill = "grey80",
  linewidth = 1,
  linecolor = "white"),)
) %>% table_cell_bg(row = c(3),column = c(2),fill= ggplot2::alpha("#79AF97FF",0.6)) %>%
      table_cell_bg(row = c(4),column = c(2),fill=ggplot2::alpha("#79AF97FF",0.6)) %>%
      table_cell_bg(row = c(4),column = c(3),fill=ggplot2::alpha("#79AF97FF",0.6))  %>%
      table_cell_bg(row = c(2),column = c(3),fill=ggplot2::alpha("#B24745FF",0.6)) %>%
      table_cell_bg(row = c(2),column = c(4),fill=ggplot2::alpha("#B24745FF",0.6)) %>%
      table_cell_bg(row = c(3),column = c(4),fill=ggplot2::alpha("#B24745FF",0.6)) -> x
cairo_pdf(file = "./3_output_figure/F1B_2_HR.pdf",width = 6,height = 5)
x$theme$text$family="TNR"
print(x)
dev.off()
```
## 7.2  valid_set
```{r}
#OS
########## calculate
SIIvsSI = coxph(Surv(OS.time,OS)~valid_set_761,sur_data_30[sur_data_30$valid_set_761!="SIII",]) %>% summary()
SIIIvsSI = coxph(Surv(OS.time,OS)~valid_set_761,sur_data_30[sur_data_30$valid_set_761!="SII",]) %>% summary()
SIIIvsSII = coxph(Surv(OS.time,OS)~valid_set_761,sur_data_30[sur_data_30$valid_set_761!="SI",]) %>% summary()

#########plot

data.frame(#subtypes = c("SI","SII","SIII"),
           SI = c("",SIIvsSI$logtest[3]%>% round(.,4),SIIIvsSI$logtest[3]%>% round(.,4)),
           SII = c(F.pasteHR(SIIvsSI$conf.int),"",SIIIvsSII$logtest[3]%>% round(.,4)),
           SIII = c(F.pasteHR(SIIIvsSI$conf.int),F.pasteHR(SIIIvsSII$conf.int),"")
           ) %>% 
ggtexttable(rows = c("SI","SII","SIII"),cols=c("SI","SII","SIII"),
            theme = ttheme(colnames.style=colnames_style(color = "black",face = "bold",size = 12,fill = "grey80",linewidth = 1,
  linecolor = "white"),rownames.style=rownames_style(  color = "black",
  face = "bold",
  size = 12,
  fill = "grey80",
  linewidth = 1,
  linecolor = "white"))
) %>% table_cell_bg(row = c(3),column = c(2),fill= ggplot2::alpha("#79AF97FF",0.6)) %>%
      table_cell_bg(row = c(4),column = c(2),fill=ggplot2::alpha("#79AF97FF",0.6)) %>%
      table_cell_bg(row = c(4),column = c(3),fill=ggplot2::alpha("#79AF97FF",0.6))  %>%
      table_cell_bg(row = c(2),column = c(3),fill=ggplot2::alpha("#B24745FF",0.6)) %>%
      table_cell_bg(row = c(2),column = c(4),fill=ggplot2::alpha("#B24745FF",0.6)) %>%
      table_cell_bg(row = c(3),column = c(4),fill=ggplot2::alpha("#B24745FF",0.6))-> x
cairo_pdf(file = "./3_output_figure/F1C_1_HR.pdf",width = 6,height = 5)
x$theme$text$family="TNR"
print(x)
dev.off()

#RFS
########## calculate
SIIvsSI = coxph(Surv(RFS.time,RFS)~valid_set_761,sur_data_30[sur_data_30$valid_set_761!="SIII",]) %>% summary()
SIIIvsSI = coxph(Surv(RFS.time,RFS)~valid_set_761,sur_data_30[sur_data_30$valid_set_761!="SII",]) %>% summary()
SIIIvsSII = coxph(Surv(RFS.time,RFS)~valid_set_761,sur_data_30[sur_data_30$valid_set_761!="SI",]) %>% summary()

#########plot

data.frame(#subtypes = c("SI","SII","SIII"),
           SI = c("",SIIvsSI$logtest[3]%>% round(.,4),SIIIvsSI$logtest[3]%>% round(.,4)),
           SII = c(F.pasteHR(SIIvsSI$conf.int),"",SIIIvsSII$logtest[3]%>% round(.,4)),
           SIII = c(F.pasteHR(SIIIvsSI$conf.int),F.pasteHR(SIIIvsSII$conf.int),"")
           ) %>% 
ggtexttable(rows = c("SI","SII","SIII"),cols=c("SI","SII","SIII"),
            theme = ttheme(colnames.style=colnames_style(color = "black",face = "bold",size = 12,fill = "grey80",linewidth = 1,
  linecolor = "white"),rownames.style=rownames_style(  color = "black",
  face = "bold",
  size = 12,
  fill = "grey80",
  linewidth = 1,
  linecolor = "white"))
) %>% table_cell_bg(row = c(3),column = c(2),fill= ggplot2::alpha("#79AF97FF",0.6)) %>%
      table_cell_bg(row = c(4),column = c(2),fill=ggplot2::alpha("#79AF97FF",0.6)) %>%
      table_cell_bg(row = c(4),column = c(3),fill=ggplot2::alpha("#79AF97FF",0.6))  %>%
      table_cell_bg(row = c(2),column = c(3),fill=ggplot2::alpha("#B24745FF",0.6)) %>%
      table_cell_bg(row = c(2),column = c(4),fill=ggplot2::alpha("#B24745FF",0.6)) %>%
      table_cell_bg(row = c(3),column = c(4),fill=ggplot2::alpha("#B24745FF",0.6))-> x
cairo_pdf(file = "./3_output_figure/F1C_2_HR.pdf",width = 6,height = 5)
x$theme$text$family="TNR"
print(x)
dev.off()

##SIII vs SI&SII

SIIIvsSISII = coxph(Surv(RFS.time,RFS)~valid_set_761_2,sur_data_30) %>% summary()


#########plot

data.frame(#subtypes = c("SI","SII","SIII"),
           SI = c("",SIIvsSI$logtest[3]%>% round(.,4),SIIIvsSI$logtest[3]%>% round(.,4)),
           SII = c(F.pasteHR(SIIvsSI$conf.int),"",SIIIvsSII$logtest[3]%>% round(.,4)),
           SIII = c(F.pasteHR(SIIIvsSI$conf.int),F.pasteHR(SIIIvsSII$conf.int),"")
           ) %>% 
ggtexttable(rows = c("SI","SII","SIII"),cols=c("SI","SII","SIII"),
            theme = ttheme(colnames.style=colnames_style(color = "black",face = "bold",size = 12,fill = "grey80",linewidth = 1,
  linecolor = "white"),rownames.style=rownames_style(  color = "black",
  face = "bold",
  size = 12,
  fill = "grey80",
  linewidth = 1,
  linecolor = "white"))
) %>% table_cell_bg(row = c(3),column = c(2),fill= ggplot2::alpha("#79AF97FF",0.6)) %>%
      table_cell_bg(row = c(4),column = c(2),fill=ggplot2::alpha("#79AF97FF",0.6)) %>%
      table_cell_bg(row = c(4),column = c(3),fill=ggplot2::alpha("#79AF97FF",0.6))  %>%
      table_cell_bg(row = c(2),column = c(3),fill=ggplot2::alpha("#B24745FF",0.6)) %>%
      table_cell_bg(row = c(2),column = c(4),fill=ggplot2::alpha("#B24745FF",0.6)) %>%
      table_cell_bg(row = c(3),column = c(4),fill=ggplot2::alpha("#B24745FF",0.6))-> x
cairo_pdf(file = "./3_output_figure/F1C_2_HR.pdf",width = 6,height = 5)
x$theme$text$family="TNR"
print(x)
dev.off()
```

# 8. subtype-specfic pathways of  all sample(figure1D、1E)
## 8.1 combine train-set with valid-set
```{r}
## id-trans
gene_4032 <- bitr(geneID = row.names(dataTN),fromType = "UNIPROT",toType = "ENTREZID",OrgDb = "org.Hs.eg.db")
gene_4032 <- gene_4032[!duplicated(gene_4032$ENTREZID),]
gsea.data_4032 <- dataT[gene_4032$UNIPROT,]
rownames(gsea.data_4032) <- gene_4032$ENTREZID


## kegg data-base
gmt_kegg <- readLines("./1_raw_data/1.Proteomic data/c2.cp.kegg.v7.4.entrez.gmt")
no_name <- function(gmt.list){
  as.character(unlist(strsplit(gmt.list,
                               split="\t",fixed = T
  )))[-2]
}
gmt_kegg[1]
F.gmt_fix <- function(gmt){
  database_list_gsva <- lapply(gmt, no_name)
for(i in 1:length(database_list_gsva)){
  
  names(database_list_gsva)[i] <- database_list_gsva[i][[1]][1]
  database_list_gsva[i][[1]] <- database_list_gsva[i][[1]][-1]
}
  return(database_list_gsva)
}

kegg_list <- F.gmt_fix(gmt_kegg)
```
## 8.2 ssGSEA
```{r}
ssgsea_kegg <- gsva(as.matrix(gsea.data_4032),gset.idx.list = kegg_list,method="gsva",mx.diff=FALSE,
                  kcdf="Gaussian") %>% as.data.frame()


```
## 8.3 difference analysis
```{r}
SI = which(colnames(ssgsea_kegg) %in% rownames(sur_152.data)[sur_152.data$sub761=="SI"])
SII = which(colnames(ssgsea_kegg) %in% rownames(sur_152.data)[sur_152.data$sub761=="SII"])
SIII = which(colnames(ssgsea_kegg) %in% rownames(sur_152.data)[sur_152.data$sub761=="SIII"])

ssgsea_kegg.diff <- apply(ssgsea_kegg,1,function(k){
  data.frame(SIIvsSI.fc = mean(k[SII] - k[SI]),
             SIIvsSI.p = wilcox.test(k[SII],k[SI])[[3]],
             SIIIvsSI.fc = mean(k[SIII] - k[SI]),
             SIIIvsSI.p = wilcox.test(k[SIII] - k[SI])[[3]],
             SIIIvsSII.fc = mean(k[SIII] - k[SII]),
             SIIIvsSII.p = wilcox.test(k[SIII] - k[SII])[[3]])
}) %>% do.call(rbind,.)

```
## 8.4 cluster.1
```{r}
SI_up.k <- ssgsea_kegg.diff %>% filter(SIIvsSI.fc<0&SIIIvsSI.fc<0&SIIvsSI.p<0.01&SIIIvsSI.p<0.01)
SIII_up.k <- ssgsea_kegg.diff %>% filter(SIIIvsSII.fc>0&SIIIvsSI.fc>0&SIIIvsSI.p<0.01&SIIIvsSII.p<0.01)
all_diff.k <- ssgsea_kegg.diff %>% filter(SIIIvsSII.p<0.01|SIIIvsSI.p<0.01|SIIvsSI.p<0.01)
other.k <- setdiff(rownames(all_diff.k),c(rownames(SI_up.k),rownames(SIII_up.k))) 

```
## 8.5 cluster.2
```{r}
ssgsea_kegg.scale <- t(ssgsea_kegg) %>% scale() %>% t() %>% as.data.frame()
SI_up.order <- dist(ssgsea_kegg[rownames(SI_up.k),]) %>% hclust() %>% use_series(order) %>% rownames(SI_up.k)[.]
SIII_up.order <- dist(ssgsea_kegg[rownames(SIII_up.k),]) %>% hclust() %>% use_series(order) %>% rownames(SIII_up.k)[.]
## cut other.k to  two class
cut.result <- dist(ssgsea_kegg[other.k,]) %>% hclust() %>% cutree(k=2)
other.1 <- names(cut.result[cut.result==1])
other.2 <- names(cut.result[cut.result==2])

##
other.1.order <- dist(ssgsea_kegg[other.1,]) %>% hclust() %>% use_series(order) %>% other.1[.]
other.2.order <- dist(ssgsea_kegg[other.2,]) %>% hclust() %>% use_series(order) %>% other.2[.]

kegg_cluster <- data.frame(row.names = c(SI_up.order,SIII_up.order,other.1,other.2),
                           kegg = c(SI_up.order,SIII_up.order,other.1,other.2),
                           cluster=c(rep("k1",44),rep("k4",50),rep("k2",18),rep("k3",36)))
```
## 8.6 heatmap
```{r}
p8.6.1 <- F.heatmap(data = ssgsea_kegg.scale[SI_up.order,],gene.order = SI_up.order,sample.order = colnames(ssgsea_kegg.scale)[c(SI,SII,SIII)],fillname = "Enrichment score")
p8.6.1.tree <- F.tree(dist(ssgsea_kegg[SI_up.order,]) %>% hclust())


p8.6.2 <- F.heatmap(data = ssgsea_kegg.scale[other.1.order,],gene.order = other.1.order,sample.order = colnames(ssgsea_kegg.scale)[c(SI,SII,SIII)],fillname = "Enrichment score")
p8.6.2.tree <- F.tree(dist(ssgsea_kegg[other.1.order,]) %>% hclust())

p8.6.3 <- F.heatmap(data = ssgsea_kegg.scale[other.2.order,],gene.order = other.2.order,sample.order = colnames(ssgsea_kegg.scale)[c(SI,SII,SIII)],fillname = "Enrichment score")
p8.6.3.tree <- F.tree(dist(ssgsea_kegg[other.2.order,]) %>% hclust())

p8.6.4 <- F.heatmap(data = ssgsea_kegg.scale[SIII_up.order,],gene.order = SIII_up.order,sample.order = colnames(ssgsea_kegg.scale)[c(SI,SII,SIII)],fillname = "Enrichment score")
p8.6.4.tree <- F.tree(dist(ssgsea_kegg[SIII_up.order,]) %>% hclust())

##anno 
p8.6.anno <- data.frame(sample=rownames(sur_152.data),
                       sub=sur_152.data$sub761)
ggplot(p8.6.anno,aes(x=factor(sample,levels = colnames(ssgsea_kegg.scale)[c(SI,SII,SIII)]),y="",fill=sub))+
  geom_tile()+
  theme_void()+
  scale_fill_manual(values = c("forestgreen","midnightblue","red4"))->p8.6.5

p8.6.1 %>% insert_top(p8.6.5,height = 0.08) %>%
  insert_bottom(p8.6.2,height = 0.5) %>%
  insert_bottom(p8.6.3,height = 0.8) %>%
  insert_bottom(p8.6.4,height = 1)
```
## 8.7 save plot
```{r}
cairo_pdf(filename = "./3_output_figure/F1D_1.pdf",width = 7,height = 3)
p8.6.1 %>% insert_left(p8.6.1.tree,width = 0.05)
dev.off()

cairo_pdf(filename = "./3_output_figure/F1D_2.pdf",width = 7,height = 1.5)
p8.6.2 %>% insert_left(p8.6.2.tree,width = 0.05)
dev.off()

cairo_pdf(filename = "./3_output_figure/F1D_3.pdf",width = 7,height = 2)
p8.6.3%>% insert_left(p8.6.3.tree,width = 0.05)
dev.off()

cairo_pdf(filename = "./3_output_figure/F1D_4.pdf",width = 7,height = 3)
p8.6.4%>% insert_left(p8.6.4.tree,width = 0.05)
dev.off()

cairo_pdf(filename = "./3_output_figure/F1D_5.pdf",width = 7,height = 0.5)
print(p8.6.5)
dev.off()
```
# 9. Sankey diagram
```{r}
kegg_class$cluster <- kegg_cluster[kegg_class$KEGG,"cluster"]
kegg_class_stastic <- data.frame(Var = table(kegg_class$group) %>% names()) 
table(kegg_class[kegg_class$cluster=="k1","group"]) 
table(kegg_class[kegg_class$cluster=="k2","group"])
table(kegg_class[kegg_class$cluster=="k3","group"])
table(kegg_class[kegg_class$cluster=="k4","group"])

sankey.data <- test <- data.frame(response = c("Other","Proliferation","Metabolism","Immune","Metastases","Other","Proliferation","Metabolism","Immune","Metastases",
                                "Other","Proliferation","Metabolism","Immune","Metastases",
                                "Other","Proliferation","Metabolism","Immune","Metastases"),
                   freq = c(4.5,0,95.5,0,0,
                            16.7,11.1,72.2,0,0,
                            16.7,25,16.7,36.1,5.5,
                            26,30,4,22,18),
                   subject = c(1:5,1:5,1:5,1:5),
                   survey=c(rep("k1",5),rep("k2",5),rep("k3",5),rep("k4",5))
                   )

sankey.data$survey <- factor(test$survey,levels = c("k1","k2","k3","k4"))
sankey.data$response <- factor(test$response,levels = c("Metabolism","Proliferation","Immune","Metastases","Other"))
ggplot(sankey.data,
       aes(x = survey, stratum = response, alluvium = subject,
           y = freq,
           fill = response, label = response)) +
  geom_flow(size=2,alpha=1) +
 geom_stratum(alpha = 1,linetype=5) +
  #geom_text(stat = "stratum", size = 3,family="AB") +

  theme_bw()+
  scale_x_discrete(expand = c(0.05,0))+
  scale_y_continuous(expand = c(0,0))+
  theme(
        axis.text = element_text(color = "black",family = "TNR",size = 10),
        #axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title = element_text(color = "black",family = "TNR",size = 13)
        )+
  theme(legend.position = "right")+
  #geom_segment(x=3.16,y=0,yend=100,xend=3.16,hjust=0.5)+
  #geom_segment(x=c("k3"),y=0,yend=100,xend="k3")+
  labs(x="",y="Percent of KEGG (%)")+
  scale_fill_manual(breaks = c("Metabolism","Proliferation","Immune","Metastases","Other"),values = c("#2e8b57","#E99B12","tomato3","salmon4","grey80"))+
  #geom_text(x="k1",y=85,label=c("Metabolism"),hjust=0.03,family="AB",size=6.5)+
  #geom_text(x="k3",y=52,label=c("Proliferation"),hjust=0.15,family="AB",size=5)+
  #geom_text(x="k3",y=81,label=c("Immune"),hjust=0.17,family="AB",size=6)+
  #geom_text(x="k4",y=31,label=c("Metastases"),hjust=0.9,family="AB",size=4)+
  #geom_text(x="k3",y=15,label=c("Other"),hjust=1,family="AB",size=5)+
  theme(plot.margin = unit(c(1,1,0,0),"line"),
        panel.grid = element_blank()) ->p9

cairo_pdf(filename = "./3_output_figure/F1_E.pdf",width = 7,height = 5)
print(p9)
dev.off()
```
# save
```{r}
save(list = ls(),file = "./Code of Figure 1.Rdata")
save(sub_specfics,file = "./2_output_data/761sub_specfics.Rdata")
save(sur_152.data,file = "./2_output_data/sur_152.Rdata")
load("./Code of Figure 1.Rdata")

```

